---
title: "HW3 - ANOVA"
author: "Shiloh Bradley"
date: "6/10/2020"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(ggplot2)

setwd("~/Desktop/Personal Computer/MBA/HOA 732/")
```

```{r Problem 3: Test if the average sales are equal for three promotions} 
S <- read.csv("restaurant4anova.csv",header=TRUE)
head(S)
tail(S)
```
These box plots show us that promo 3 is the best. No promotion is the worst. There is an overlap between these three though. We wil lse 
```{r Step 1 Box Plot of Sales by Promo}
BX2 <- ggplot(S,aes(x=factor(Promo),y=Sales)) + geom_boxplot()+
       stat_summary(fun = "mean", geom="point",color="red")
BX2 + xlab("Promotion")
```

```{r}
p.dot <- ggplot(S,aes(x=factor(Promo),y=Sales)) + 
  geom_dotplot(binaxis='y', stackdir='center')
p.dot + xlab("Promotion")
```

```{r}
#xbari.	65.19592	76.84204	77.82536	99.882
#xbar..	82.53757882
S$Promo <- factor(S$Promo)
p.dot <- ggplot(S,aes(x=Promo,y=Sales,fill=Promo)) + 
  geom_dotplot(binaxis='y', stackdir='center') +
  geom_hline(yintercept = c(65.19592,76.84204,77.82536,99.882), color=c("red","green","cyan  ","purple"))+
  geom_hline(yintercept = 82.54)
p.dot 
```

```{r Descriptive stats}
# summary stats
xbar <- tapply(S$Sales, S$Promo, mean)
SD <- tapply(S$Sales, S$Promo, sd)
round(xbar,2)
round(SD,2)
```
Sales is the response variable. Promo is a categorical predictor of Sales.
```{r use 1-way ANOVA to compare mean sales of 3 promos}
a1 <- aov(data=S,Sales~factor(Promo))
summary(a1) ## Shows us the ANOVA table
names(a1)
``` 

Since P-Value is less than .05, we reject H0 and conclude that not all means are equal.
Q-Q plot shows that: residuals plot along the normal line, suggesting residuals are normal.
In addition, P-value of Shapiro-Wilks' normality test is > 0.05, so normality is verified. This makes the P-value from F-test accurate. 
```{r normality test for residuals from ANOVA}
#qq plot with normal line (normality test)
df <- as.data.frame(a1$residuals)
colnames(df)[1] <- "residuals"

shapiro.test(df$residuals) # W = 0.97087, p-value = 0.05123

ggplot(df)+stat_qq(aes(sample=residuals)) + 
  geom_qq_line(aes(sample=residuals))+
  geom_text(aes(x=0.5, y=-20, label="Shapiro-test p-value = 0.05123"))
```